---
title: "Hate crimes"
author: "Qiqi Liang, Micheal Cinnamon, Jiyoon Kim"
date: "11/14/2022"
output: html_document
---

## Dynamic Paragraph

'''
From summary statistics, the number of observations in our hate crime dataset is 219577. We have 28 variables and the year
with the max occurances (or recorded observations) of hate crimes was the 2020. We also found the state with the max crime
rates during this year was California and the max cause for the hate crime was "anti-blackness." 
'''

## Table of aggregated data: 

```{r, echo=FALSE}
hatecrime <- read.csv("https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/hate_crime.csv")
hatecrime <- hatecrime %>% group_by(BIAS_DESC)

State <- c(hatecrime$STATE_NAME)
Bias <- c(hatecrime$BIAS_DESC)

Crime <- data.frame(State, Bias)

Crime <- group_by(Crime, Crime$Bias) %>% filter(n() > 0)

Table <- table(as.character(Crime$State, Crime$Bias))

View(Table)
```

Above is an aggregated table representing the amount of hate crimes that happened or were documented in each state.
From this table, we can see that the maximum frequency of hate crime in 2020 occured in California while the lowest
frequency of hate crimes (with the exception of Guam) occured in Wyoming.

## Including Plots

```{r, echo=FALSE}
install.packages("ggplot2")
install.packages("plotly")
install.packages("tidyverse")
library("dplyr")
library("tidyr")
library("ggplot2") 
library("plotly")
library("tidyverse")
```

```{r, echo=FALSE}
# Chart one

hate_crime <- read.csv(file = 'https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/hate_crime.csv') %>%
  rename(year = DATA_YEAR,
         state = STATE_NAME,
         state_po = STATE_ABBR) %>%
  filter(year == 2020)

political_party <- read.csv(file = 'https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/political_party.csv') %>%
  rename(party = party_simplified) %>%
  mutate(state = str_to_title(state),
         party = str_to_title(party)) %>%
  mutate(state = recode(state, `District Of Columbia` = "District of Columbia")) %>%
  filter(year == 2020)

us_population_2020 <- read.csv(file = 'https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/US_Population_2020.csv') %>%
  rename(state = State,
         population = X2020.Census) %>%
  select(state, population)

#how many hate crimes in each state
state_crimes <- hate_crime %>%
  group_by(state) %>%
  summarize(cases = n()) %>%
  left_join(us_population_2020, by = "state") %>%
  mutate(crime_ratio = cases / population) %>%
  select(state, crime_ratio)

#calculating party affiliation by state
state_by_votes_2020 <- political_party %>%
  group_by(state) %>%
  filter(candidatevotes == max(candidatevotes, na.rm = TRUE)) %>%
  select(state, party)

#join dataset to create chart
party_crimerates <- left_join(state_by_votes_2020, state_crimes, by = "state")


# Chart
politics_crime_chart <- ggplot(data = party_crimerates, 
                               aes(x = party, y = state, fill = crime_ratio)) + 
  geom_tile(color = "white") +
  labs(
    title = "Hate Crime Rates and Party Affiliation by State") +
  theme(axis.text = element_text(size = 5), 
        axis.title = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        plot.title = element_text(size = 8),
        legend.key.size = unit(0.4, 'cm'),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 7))

View(politics_crime_chart)

```

This chart above is a heatmap chart to answer the research question: "How does the political party affilitation by state
correspond with crime rates?". This reveals information regarding which party each state voted for presidency in the 2020 election.
It also calculates hate crime rates in 2020 by each state by calculating the number of crimes in each state divided by the 
population in each state. By looking at the chart, we can see that the states that voted for the Republican party is darker, while
the states that voted for the Democrats are generally lighter in colour (less hate crime).

```{r, echo=FALSE}
# Chart two

# Load hate crimes dataset. Filter year to 2020

hate_crimes <- read.csv("https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/hate_crime.csv") %>%
  rename(year = DATA_YEAR) %>%
  filter(year == 2020)

# Group and summarise data based on motivations/biases.
# The rows for some of the columns look like this: "Anti-Asian;Anti-Hispanic;Anti-Black"
# We need to separate this to get an accurate count of biases
biases_grouped_df <- hate_crimes %>%
  group_by(BIAS_DESC) %>%
  summarise(num_biases = n())

# Split the BIAS_DESC column into a list -> "[Anti-Asian, Anti-Hispanic, Anti-Black]"
# using a semi-colon delimeter
# "Unnest" to make each element in the list its own row
biases_grouped_df_splited <- biases_grouped_df %>%
  mutate(biases_array = strsplit(BIAS_DESC, ";")) %>%
  unnest(biases_array)

# Count unique biases and store in new dataframe to plot
final_biases_df <- biases_grouped_df_splited %>%
  group_by(biases_array)%>%
  summarise(new_num_biases = n())

# Convert to percentages

final_biases_df <- final_biases_df %>%
  mutate(proportion = (new_num_biases/sum(new_num_biases))*100) %>%
  mutate(year = "2020")

# Simple bar plot
fig <- plot_ly(data = final_biases_df, x = ~biases_array, y = ~proportion, type = 'bar') %>%
  layout(xaxis = list(autotypenumbers = 'strict', title = 'Types of biases'),
         yaxis = list(title = 'Proportion'),
         plot_bgcolor='#e5ecf6',
         xaxis = list(
           zerolinecolor = '#ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'),
         yaxis = list(
           zerolinecolor = '#ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'))

View(fig)

```

This chart is a bar plot that attempts to answer the question "What are the causes and motivations behind hate crimes?"
There are 35 unique biases included in this dataset, for example "Anti-Black" or "Anti-Hispanic." Since some of the rows included
multiple biases, a new dataframe was created to "unnest" the data and each bias was then summarized. From the bar plot, "Anti-Black"
was the most frequent cause of hate crimes in 2020, followed by "Anti-Gay (male)" and "Anti-Jewish." Additionally, "Anti-Asian"
and "Anti-Hispanic" were also a large reported motivation behind hate crimes. In general, it seems that racially motivated hate crimes
were most common.

```{r, echo=FALSE}
# Chart third

hatecrimelaws <- read.csv("https://raw.githubusercontent.com/info201a-au2022/project-group-3-section-ah/main/data/Hate_crime_laws_by_state.csv")

Yes <- filter(hatecrimelaws, Race.Religion.Ethnicity == "YES")
No <- filter(hatecrimelaws, Race.Religion.Ethnicity == "NO")

States <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")
WithLaws <- c(hatecrimelaws$Race.Religion.Ethnicity)

withlaws1 <- WithLaws

withlaws1 <- as.character(withlaws1)
withlaws1[withlaws1 == "YES"] <- 1     
withlaws1[withlaws1 == "NO"] <- 0      
Has_A_Law <- as.numeric(withlaws1)

Laws <- data.frame(States, Has_A_Law)

plot <- ggplot(data = Laws, aes(x = States, y = Has_A_Law)) + geom_bar(stat = "identity") 
View(plot)
```

The third chart above is a bar chart that attempts to answer the research question, "How are hate crimes punishable across different states?"
From this chart, we can see that Arizona, Indiana, South Carolina, and Wyoming do not have laws that punish hate crimes.  